---
title: "ITEC 621 Exercise 2 - Foundations"
author: "Dawoon (Kate) Jung"
date: "1/28/2021"
output:
  word_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
subtitle: Descriptive and Predictive Analytics
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=T, warning=F, message=F)
```

## General Instructions

In this exercise you will do quick descriptive and predictive analytics to evaluate if the Salaries data set (with professor salaries) supports the **gender pay gap hypothesis**. 

First, download the R Markdown template for this exercise **Ex1_Foundations_YourLastName.Rmd** and save it with your own last name **exactly**. Then  open it in R Studio and complete all the exercises and answer the questions below in the template. Run the code to ensure everything is working fine. When done, upload onto blackboard, knit your R Markdown file into a Word document and upload it into Blackboard. If for some reason you can't knit a Word file, knit an HTML file and save it as a PDF. Blackboard will not accept HTML files, but will take your PDF.

## 1. Descriptive Analytics

**1.1 Examine the data**

Is there a gender pay gap? Let's analyze this important question using professor salaries.

Load the library **{car}**, which contains the **Salaries** data set. Then, list the first few records with `head(Salaries)`. The display the `summmary()` for this dataset, which will show frequencies. 

Then, load the library **{psych}** which contains the `describe()` function and use this function to list the descriptive statistics for the data set. 

Then display the median salary grouped by gender using the `aggregate()` function (feed grouping variables, dataset and aggregate function, i.e., `salary ~ sex, Salaries, mean`)

```{r GenderGap}
library(car)
head(Salaries)
library(psych)
describe(Salaries)
aggregate(salary ~ sex, Salaries, mean)

```

**1.2 Correlation, Boxplots and ANOVA**

Load the library **GGally** and run the **ggpairs()** function on the **salary** (notice that the **Salary** data set is capitalized, whereas the variable **salary** is not), **sex** and **yrs.since.phd** variables (only) in the **Salaries** data set to display some basic descriptives and correlation visually. Please label your variables appropriately (see graph below).

Tips: `ggpairs()` requires a **data frame**. So you need to use the `data.frame()` function to bind the necessary column vectors into a data frame (e.g., `ggpairs(data.frame("Salary"=Salaries$salary, etc.)`. Notice the difference in the quality of the graphics and how categorical variables are labeled. Also, add the attribute `upper=list(combo='box')` at the end to get labels for the boxplot.

Finally, conduct an ANOVA test to evaluate if there is a significant difference between mean salaries for male and female faculty. Feed `Salaries$salary ~ Salaries$sex` into the `aov()` function. Embed the `aov()` function inside the `summary()` function to see the statistical test results. 

```{r Descriptives}
library(GGally)
attach(Salaries)

ggpairs(data.frame("Salary" = Salaries$salary, "Gender" = Salaries$sex, "Years" = Salaries$yrs.since.phd), upper=list(combo='box'))

summary(aov(Salaries$salary ~ Salaries$sex))

```

**1.3 Preliminary Interpretation**

Based on the output above, does it appear to be a gender pay gap? Why or why not. In your answer, please refer to as much of the data above to support your answer.

```{r}
# ANOVA shows that there is a statistically significant difference in salary across gender. However, in the boxplot of salary and gender, since the two boxes are largely overlapped, it does not support their differences. 

# There is another variable that has an association with salary is years since they achieved a Ph.D. degree. The correlation between years and salary is about 0.419, which indicated the two variables are highly correlated. The barplot of Years and Gender shows that the years of Ph.D. of male faculty are higher than the female faculty's years. Therefore, it can be one possible explanation of the gender pay gap.

```



## 2. Basic Predictive Modeling

**2.1 Salary Gender Gap: Simple OLS Regression**

Suppose that you hypothesize that there is a salary gender pay gap. Fit a linear model function `lm()` to test this hypothesis by predicting salary using only **sex** as a predictor. Store the results in an object called `lm.fit.1`, then inspect the results using the `summary()` function. Do these results support the salary gender gap hypothesis? Briefly explain why.

```{r lmfit1}

lm.fit.1 <- lm(salary ~ sex, data = Salaries)
summary(lm.fit.1)

# Yes. The linear regression model shows that on average, when the faculty is male, the salary is higher than female faculty by about $14,088. 
```

**2.2 Multivariate OLS Regression**

Now fit a linear model with **sex** and **yrs.since.phd** as predictors and save it in an object named `lm.fit.2`. Then inspect the results  using the `summary()` function. Do these results support the salary gender gap hypothesis? Briefly explain why.

```{r lmfit2}

lm.fit.2 <- lm(salary ~ sex + yrs.since.phd, data = Salaries)
summary(lm.fit.2)

# The evidence is not conclusive. On average, male faculty get paid $7923.6 more than female faculty, holding the years since achieving a Ph.D. degree. However, the p-value is only 0.0915; this effect is only moderately significant.

```

**2.3 Comparing Models with ANOVA F-Test**

Run an ANOVA test using the `anova()` funtion to compare **lm.fit.1** to **lm.fit.2**.

```{r anova}



anova(lm.fit.1)

anova(lm.fit.2)



```

**2.4 Interpretation**

Provide your brief conclusions (in no **more than 3 lines**) about whether you think there is a gender pay gap based on this analysis (you will expand this analysis much further in HW2). First, which `lm()` model is better and why? Then, compare the best predictive model of the two against the descriptive analytics results you obtained in section 1 above. If the null hypothesis is that there is no gender pay gap, is this hypothesis supported? Why or why not?



```{r}
# With lm.fit.2 model, the null hypothesis is failed to reject due to the lack of significance. In order to reject the null hypothesis, the rule of thumb number of a coefficient is p < 0.05. On the other hand, the lm.fit.1 model rejects the hull hypothesis with statistical significance. It indicates that the predictor yrs.since.phd correlates with the other predictor, gender. As the ANOVA test shows yrs.since.phd is an important prediction revealing the gender pay gap. 
```

